Abstract

### 2025-10-06 output/Main_text.Rmd
Match table footnote style to figure footnote's
Define latent-space test transformation

We make a few observations here.
First, the transformation to the orthonormal bases of the two spaces allows
mapping the $`r MODEL_ACRONYM`$ parameters interchangeably in the two spaces,
by projecting one space onto the other.
This allows representing the coordinates of the discrimination vectors
in the latent space and, conversely,
the coordinates of the latent trait vectors in the test space.
As we can see, the invariance of the model holds,
no matter in which space we represent the coordinates.
Second, and most importantly,
given that the test space is defined as the set of item parameters
(see section *Test Space Definition*),
then they should be represented in this space, and not in the latent space.
Therefore, we may find the direction angles of $`r DISCR_VECTOR`$
in the test space:
Applying Proposition 1, if $`r ANGLE_VEC_TS_ITEM`$ is
the vector of direction angles of $`r DISCR_VECTOR`$,
in the test space $`r TEST_SPACE`$ we get

@@

::: {custom-style="FigureNote"}
*Note.* The items are represented along with
a standard bivariate normal distribution, with null correlation (a),
or with correlation $`r CORR_OBL_OUT`$ (b-d).
The contour plots represent the `r contour_breaks_out` of the maximum density,
from outer to inner, respectively.
:::

>>>>

We make a few observations here.
First, the transformation to the orthonormal bases of the two spaces allows
mapping the $`r MODEL_ACRONYM`$ parameters interchangeably in the two spaces,
by projecting one space onto the other;
that is, we can set

\begin{equation}
  `r DISCR_COORDS_TRANSF_EQ`                  (\#eq:trait-coords-orth-spaces-eq)
\end{equation}

for any point, $`r TRAIT_VECTOR`$ or $`r DISCR_VECTOR`$.
This allows representing the coordinates of the discrimination vectors
in the latent space and, conversely,
the coordinates of the latent trait vectors in the test space:
By \@ref(eq:change-basis), \@ref(eq:discr-vector-std),
and  \@ref(eq:trait-coords-orth-spaces-eq)
any coordinate vector $`r DISCR_COORDS`$ in $`r TEST_SPACE_BASIS`$
can be expressed in
$`r LS_BASIS`$ as $`r DISCR_COORDS_TS_EQ`$, and vice versa.
As we can see, the invariance of the model holds,
regardless of the space in which we represent the coordinates.

Second, and most importantly,
given that the test space is defined as the set of item parameters
(see section *Test Space Definition*),
then they should be represented in this space, and not in the latent space.
Therefore, we may find the direction angles of $`r DISCR_VECTOR`$
in the test space:
Applying Proposition 1, if $`r ANGLE_VEC_TS_ITEM`$ is
the vector of direction angles of $`r DISCR_VECTOR`$,
in the test space $`r TEST_SPACE`$ we get

@@

::: {custom-style="FigureNote"}
*Note.* The items are represented along with
a standard bivariate normal distribution, with null correlation (a),
or with correlation $`r CORR_OBL_OUT`$ (b-d).
The contour plots represent (from outer to inner)
the `r contour_breaks_out` of the maximum density.
:::
